---
title: "Biostats-Sample"
author: "Alexis Garduno"
date: '2023-08-04'
output:
  pdf_document: default
  html_document: default
---
##Description: To perform exact matching on a select group of subjects with a healthcare encounter.
##Input: Encounter_Table, Patient_Table
##Output: Median Age (years) at Index for Exposed Cohort; Aggregate Time at Risk for Exhosed Cohort (person-days)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
```

##Load tables
```{r cars}
path <- 'C:/Users/Alexis Garduno/iCloudDrive/Documents/2022-2023/Personal/Interview/Graticule/biostats-sample'
encounter_table <- read.csv(paste0(path,'/encounter_table.csv'))
patient_table <- read.csv(paste0(path,'/patient_table.csv'))

summary(encounter_table)
summary(patient_table)

encounter_table <- encounter_table %>% mutate(across(ends_with("date"), ~as.Date(.x, format = "%Y-%m-%d")))
encounter_table$exposure <- as.numeric(as.logical(encounter_table$exposure))
encounter_table$outcome <- as.numeric(as.logical(encounter_table$outcome))
encounter_table$patientid <- as.factor(encounter_table$patientid)

patient_table <- patient_table %>% mutate(across(ends_with("date"), ~ as.Date(.x, format = "%Y-%m-%d")))
patient_table[,c('patientid','gender')] <- lapply(patient_table[,c('patientid','gender')] , factor)


summary(encounter_table)
summary(patient_table)
head(encounter_table)
head(patient_table)
```
##Merge the dataset and check range
```{r cars}
#merged datasets; checked that all ids matched between datasets
dim(patient_table); dim(encounter_table)
patient_table2 <- merge(patient_table,encounter_table,by='patientid',all=FALSE)
#calculate age
patient_table2$age <- round(as.numeric(patient_table2$encounterdate - patient_table2$birthdate)/365.25,0) 
dim(patient_table2)
```

##Exact Matching
##For initialization, begin matching on the second week of data from July 1st to ensure that we have observation on participants showing no prior exposure.
```{r cars}
##less than one year of data; identify week using base package (assuming ISO week is not needed)
patient_table2 <- patient_table2 %>% mutate(encounter_wk=strftime(encounterdate,format = "%W"))

#examine count of exposed and unexposed
table(patient_table2$exposure)

#generate a dataframe with list of exposed subjects
patient_table2_exp <- patient_table2[patient_table2$exposure==1,]
#identify the earliest exposure date for participants
patient_table2_exp2 <-  patient_table2_exp %>% group_by(patientid) %>% summarise_at(vars(encounterdate), list(min = min))

#unexposed records
patient_table2_noexp <- patient_table2[patient_table2$exposure==0,]

#for unexposed records, merge back the date of exposure
copy_exp <- patient_table2_exp2
colnames(copy_exp) <- c('patientid','exp_encounterdate')
#merge back and filter
patient_table2_noexp2 <- merge(patient_table2_noexp,copy_exp,by='patientid',all.x=TRUE)
dim(patient_table2_noexp2)
#remove hospital encounters where exposed prior to unexposed record
patient_table2_noexp2 <- patient_table2_noexp2[patient_table2_noexp2$encounterdate < patient_table2_noexp2$exp_encounterdate,]
dim(patient_table2_noexp2)

#empty arrays for exposed, unexposed
exposed_arr <- c()
unexposed_arr <- c()
i=0
while(nrow(patient_table2_exp) != 0) {
 exposed <- patient_table2_exp[1,]
 unexposed <- merge(exposed,patient_table2_noexp2,by=c('encounter_wk','gender'))
 #keep first unexposed person
 if (nrow(unexposed) > 0) {
   unexposed_select <- unexposed[1,]
   unexposed_name <- unexposed_select$patientid.y
   #remove their records from the unexposed dataset
   patient_table2_noexp2 <- subset(patient_table2_noexp2,patientid!=unexposed_name)
   print('Total Number of Unexposed Records Remaining')
   print(dim(patient_table2_noexp2))
   #remove individual from the exposed group
   exposed_name <- exposed$patientid
   patient_table2_exp <- subset(patient_table2_exp,patientid!=exposed_name)
   print('Total Number of Exposed Recordss Remaining')
   print(dim(patient_table2_exp))
   #stack dataframe
   if (i==0) {
      exposed_arr <- exposed
      unexposed_arr <- unexposed_select
   } else {
      exposed_arr <- rbind(exposed_arr,exposed)
      unexposed_arr <- rbind(unexposed_arr,unexposed_select)
   }
   i = i+1
 } else {
   patient_table2_exp <- patient_table2_exp[-c(1),]
 }
 rm(exposed);rm(unexposed)
}

```
#Verify Deduplicated Sample
```{r cars}
any(duplicated(exposed_arr$patientid))
any(duplicated(unexposed_arr$patientid))
dim(exposed_arr)
#only 23 were matched based on the given criteria
```

#Median Age at Index in Exposed Cohort

The median age at index of the exposed cohort is August 6th 2022; the 95% CI for this date is shown 2023-07-21 to 2021-08-21.

```{r cars}
#calculate median index date
print('Median Index Date')
med <- median(exposed_arr$encounterdate)
med
#Standard Deviation
sd <- sd(exposed_arr$encounterdate) #days?
#Assume that this is the total number of days
upper_95ci <- med + 1.96*sd
lower_95ci <- med - 1.96*sd
paste0(paste0('95%CI: ',upper_95ci,' to ', lower_95ci))
```

#Total Time at Risk for Exposed Cohort (Person-Days)

The total time at risk for the exposed cohort (n=23) is 4910 days.

We will assume that an absence of a death record means that they are still under observation.
```{r cars}
study_end <- rep(as.Date('2022/12/31'),dim(exposed_arr)[1])
exposed_arr2 <- cbind(exposed_arr,study_end)
end_obs <- coalesce(exposed_arr2$deceaseddate,exposed_arr2$study_end)
exposed_arr2 <- cbind(exposed_arr2,end_obs) #none of the dates are before dec31st
#time at risk = index to outcome or death
#calculate days from the study index date to study end; assuming no outcome from index date to end of study period in this cohort
exposed_arr2$time_atrisk <- exposed_arr2$study_end - exposed_arr2$encounterdate
##total person-time at-risk
total_atrisk <- sum(exposed_arr2$time_atrisk)
total_atrisk
```


